---
title: "S6 Preparation of Roads and Buildings Data"
author: "Author Blank for Double Blind Peer Review"
output:
  html_document:
    df_print: paged
---

## Introduction

In this file I prepare the data *coauthor's name obfuscated* calculated with ArcMap that quantify for each plot the distance from the plot centre to each of the nearest small road, big road and building.
Below I prepare these data for use as explanatory variable in our models for ant occurence.

Input Data :
Roads: `fgeolib.geolib.swissTLM_1_2_STRASSEN\fgeolib.geolib.TLM_STRASSE_2016`

Buildings: `fgeolib.geolib.swissTLM_1_2_BAUTEN\fgeolib.geolib.TLM_GEBAEUDE_FOOTPRINT_2016` 

NFI Plot Coordinates: `Richtige Koordinaten der besuchten Probeflächen_2009-2017_Rohdaten.xlsx`

## R packages used:
In this data preparation I make use of the [tidyverse](https://www.tidyverse.org/) suite of packages.
I also use `readxl` for reading data into R from Excel files.

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
```

## Read in the Data
`Plotcent_09_17_roads_buildings.xlsx` has column names in the first row and column units in the second row.

Reading in the column names and units:
```{r}
rdbd.cn.cu.tb <- read_excel(path = '~/rwa/data/roads_buildings/Plotcent_09_17_roads_buildings_20180326.xlsx', sheet = 'Plotcent_09_17_nearest', n_max = 1)
```

Abbreviate the column names:
  abberviate 'big' as 'l' (for large) so we can reserve 'b' for 'building'
```{r rows.print=31}
CN.tb <- tibble(Orig = colnames(rdbd.cn.cu.tb)) %>%
  mutate(Abbrv = str_replace_all(string = Orig, pattern = 'near500_', replacement = '') %>%
    str_replace_all(string = ., pattern = 'bigroads', replacement = 'lr') %>%
    str_replace_all(string = ., pattern = 'smroads', replacement = 'sr') %>%
    str_replace_all(string = ., pattern = 'dist', replacement = 'dst') %>%
    str_replace_all(string = ., pattern = 'build', replacement = 'b')
  )
```

Read in the data:
```{r}
rdbd.all.tb <- read_excel(path = '~/rwa/data/roads_buildings/Plotcent_09_17_roads_buildings_20180326.xlsx', sheet = 'Plotcent_09_17_nearest', col_names = CN.tb$Abbrv, skip = 2)
```

## Subset the Data

Subset to the covariates and the plot identifier `CLNR`.
Note: FID = file-internal identifiers of the roads or buildings that were found to be "nearest" to the plots so we don't need to consider these as covariates

near_dst_b = Distance to nearest Building
near_dst_sr = Distance to nearest Small Road
near_dst_lr = Distance to nearest Large Road

```{r}
rdbd.tb <- select(rdbd.all.tb, CLNR, near_dst_b, near_dst_sr, near_dst_lr)

slice(rdbd.tb, 1:10)
```

Load the prepared ants data and subset the prepared roads and buildings data to the plots (CLNR) that were searched for ant mounds:

```{r}
load('~/rwa/data/ants/Preparations_of_Ameisen_Data_2009-2017_pourWSL_22.03.2018_xlsx/Red_Wood_Ant_Mound_Occurence.RData')
```

```{r}
rdbd.tb <- select(RWAMO.tb, CLNR) %>%
  left_join(x = ., y = rdbd.tb, by = 'CLNR')
```

Check for missing data:
```{r}
anyNA(rdbd.tb)
```

View the summary statistics for the prepared data:
```{r}
summary(rdbd.tb)
```

Write out the prepared data:
```{r, eval = FALSE}
save(list = 'rdbd.tb', file = '~/rwa/data/roads_buildings/roads_buildings_data_prepared_unfiltered.RData')
```
